---
title: How I built this blog using R, Blogdown and Hugo
author: "Felix"
date: '2017-12-23'
categories:
  - Blogging
tags:
  - how to
  - blog
  - blogdown
  - rmarkdown
  - github
metaAlignment: center
output: html_document
banner: "img/Articles/blog.jpg"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```




Material
=======
Used :

*https://tclavelle.github.io/blog/blogdown_github/

*https://bookdown.org/yihui/blogdown/github-pages.html

*https://talesofr.wordpress.com/2017/08/18/starting-a-rmarkdown-blog-with-bookdown-hugo-github/

*https://yihui.name/en/2017/08/hugo-post-summary/#fnref:I-broke-the-4th 



How to
======
1. create 2 repo
  *username/source
  *username/username.github.io
2. clone repo on computer and remove readme files
3. start new R project
4. project option -> website, select source as base directory
5. load necessary libraries

```{r librairies, message=F}
# Load blogdown package and install Hugo
library(blogdown)
install_hugo()
```

6. choose a theme
7. set "source" as working directory
8. create site
```{r}
# Create new site in our recently cloned blogdown repository
new_site(theme = "kakawait/hugo-tranquilpeak-theme",
         format = 'toml')
```

9. add folowing lines in  config.toml and update the url
```{r, eval=F}
publishDir = "X:/.../username.github.io"
themesdir = "X:/.../source/themes"
baseURL = "https://username.github.io/"

```

10. build the site
```{r, eval=F}
build_site()
```


11. in Terminal (git code), type 
```{git }
cd X:/.../username.github.io

git add *

git commit -m "update ID"

git push origin master
```

12. go to https://username.github.io/ in browser 


Update
=======
###Serving site

there is a bug when serving site, to solve it, I use the following code
```{r, eval=F, eval=F}
install.packages(c("processx", "later"))
options(blogdown.generator.server = TRUE)
```

###Summary is way to long !


update layout/Summary.html to optimize Summary readability:

since `<!--more-->` does not work with Rmarkdown + Blogdown + Hugo, I modified the theme file. 
{{.Summary}} became
```{r, eval=F}
{{ if gt (len .Summary) 200 }}{{slicestr .Summary 0 200}} {{else}} {{.Summary}} {{end}} <a href="{{ .Permalink }}" class="postShorten-excerpt_link link">[...] continue reading </a>

```
source : https://discourse.gohugo.io/t/limit-the-summary/4103/8
        

