---
title: Opitimizing Mario Kart 8 Deluxe Gameplay
author: Félix
date: '2018-02-14'
categories:
  - Knowledge Discovery
slug: opitimizing-mario-kart-8-deluxe-gameplay
---


#Import data
```{r message=FALSE, warning=FALSE}
library(readxl)
Char<- read_xlsx("Mario Kart 8 Deluxe Stat.xlsx",sheet = "Char",col_names = T, col_types = c("text",rep("numeric", 12)))

Kart <- read_xlsx("Mario Kart 8 Deluxe Stat.xlsx",sheet = "Kart",col_names = T, col_types = c("text",rep("numeric", 12)))

Tire <- read_xlsx("Mario Kart 8 Deluxe Stat.xlsx",sheet = "Tire",col_names = T, col_types = c("text",rep("numeric", 12)))

Glide <- read_xlsx("Mario Kart 8 Deluxe Stat.xlsx",sheet = "Glide",col_names = T, col_types = c("text",rep("numeric", 12)))


library(tidyverse)
Char<-Char %>% add_column(Type="Char", .after='ID')
Kart <- Kart %>% add_column(Type="Kart", .after='ID')
Tire <- Tire %>% add_column(Type="Tire", .after='ID')
Glide <- Glide %>% add_column(Type="Glide", .after='ID')

data <-  full_join(Char, Kart) %>% full_join(Glide) %>% full_join(Tire)
```


#compute mean speed and handling
```{r}
data <- data %>%mutate(Speed_mean=rowMeans(data[grepl("Speed", names(data))])) %>% mutate(Handling_mean=rowMeans(data[grepl("Handling", names(data))]))
```

```{r}
library(qtlcharts)
iplotCorr(data %>% filter(Type=="Char") %>% select(-ID, -Type, -Speed_Land:-Speed_Gliding,-Handling_Land:-Handling_Gliding))

```

#How characters are distributed between Acceleration and Speed
```{r}
ggplot(data %>% filter(Type=="Char"), aes(x=Speed_mean,y=Accel))+geom_point(aes(size=Weight, fill=`M-turbo`), pch=21)+ggrepel::geom_text_repel(aes(label=ID), min.segment.length = 0.1,point.padding = unit(1.2, 'lines'),box.padding = unit(0.6, 'lines'), force=2)+viridis::scale_fill_viridis()
```
Fastest characters are also the heaviest and the ones with lowest acceleration, handling and Mini-turbo


#Wich Kart provide the best bonus Speed/Accelaration ?
```{r}
ggplot(data %>% filter(Type=="Kart"), aes(x=Speed_mean,y=Accel))+geom_point(aes(size=Weight, fill=`M-turbo`), pch=21)+ggrepel::geom_text_repel(aes(label=ID), min.segment.length = 0.1,point.padding = unit(1.2, 'lines'),box.padding = unit(0.6, 'lines'), force=2)+viridis::scale_fill_viridis()+geom_rect(aes(ymin=-Inf, ymax=0, xmin=-Inf, xmax=0), alpha=0.01, fill="red")+geom_rect(aes(ymin=0, ymax=Inf, xmin=0, xmax=Inf), alpha=0.01, fill="green")
```


**A Speed/acceleration Bonus Ratio would be good to condensate information**


```{r}
data <- data %>% mutate("Bonus_Rate"=ifelse(Speed_mean > Accel, abs(Speed_mean/Accel),abs(Accel/Speed_mean)))

ggplot(data %>% filter(Type=="Kart"), aes(Bonus_Rate, ID))+geom_point()+ggrepel::geom_text_repel(aes(label=paste(Speed_mean,"/",Accel)), min.segment.length = 0.1,point.padding = unit(1.2, 'lines'),box.padding = unit(0.6, 'lines'), force=2)

```
W 25 Silver Arrow and Cat Cruiser provides the most bonus regarding to malus (4 times more bonus than malus)

```{r}
data$total <- rowSums(data[,c(7,8,13:17)], na.rm = T)

data %>%group_by(Type) %>%filter(total==max(total, na.rm=T))

```

#+generate all combinations
```{r}
#method1
library(gtools)

comb <- as_data_frame(permutations(n=length(data$ID),r=4,v=data$ID, repeats.allowed = F)) %>% filter(V1 %in% Char$ID & V2 %in% Kart$ID & V3 %in% Tire$ID & V4 %in% Glide$ID)

names(comb)[1] <- "ID"
comb1 <- inner_join(comb, data, by="ID")

names(comb)[1] <- "V1"
names(comb)[2] <- "ID"
comb2 <- inner_join(comb, data, by="ID")

names(comb)[2] <- "V2"
names(comb)[3] <- "ID"
comb3 <- inner_join(comb, data, by="ID")

names(comb)[3] <- "V3"
names(comb)[4] <- "ID"
comb4 <- inner_join(comb, data, by="ID")

data_combo <- comb1[,6:21]+comb2[,6:21]+comb3[,6:21]+comb4[,6:21]

data_combo <- cbind(comb[,1:4], data_combo)

# Rename a column in R
names(data_combo)[1]<-"Char"
names(data_combo)[2]<-"Kart"
names(data_combo)[3]<-"Tire"
names(data_combo)[4]<-"Glidder"
```

```{r eval=FALSE}
#method 2 (not working)
data_combo <- as.data.frame((data %>% group_by(Type) %>% sample_n(1) %>% ungroup() %>% mutate_if(is.numeric,sum) %>% ungroup() %>% mutate(ID=paste(ID, collapse=" + ")) %>% select(-Type)%>%unique()))

test <- replicate(5, { rbind(data_combo, (data %>% group_by(Type) %>% sample_n(1) %>% ungroup() %>% mutate_if(is.numeric,sum) %>% ungroup() %>% mutate(ID=paste(ID, collapse=" + ")) %>% select(-Type)%>%unique()))} )
```


```{r}

data_combo <- data_combo %>% mutate(Speed_Land2=2*Speed_Land)

data_combo <- data_combo %>% mutate(Handling_Land2=2*Handling_Land)


data_combo  <- data_combo  %>%mutate(Speed_mean=rowMeans(data_combo[,c(6:8,21)])) %>% mutate(Handling_mean=rowMeans(data_combo[,c(12:14,22)]))


data_combo$total <- rowSums(data_combo[,c(9,15:18)], na.rm = T)

data_combo$total_A_S <- rowSums(data_combo[,c(9,17)])

data_combo <- data_combo %>% mutate(W_category = cut(Weight,breaks = c(0,2,3,4,5,6),labels = c("Super Light","Light","Medium","Heavy", "Super Heavy")))



```


```{r}
library(skimr)
library(ggridges)
skim(data_combo)

ggplot(data_combo, aes(total, W_category))+geom_density_ridges(aes(fill=W_category),panel_scaling = F, alpha=0.8,rel_min_height = 0.000001)+scale_fill_brewer()
```
Lighest classes have higher totals once weight is put out of the balance

```{r}
data_combo %>% group_by(W_category)%>% summarise_if(is.numeric,mean) %>% select(W_category,Weight, Accel, Speed_mean, Handling_mean, `M-turbo`,Traction)

summary <- data_combo %>% group_by(W_category) %>% select(-Char,-Glidder, -Kart, -Tire) %>% select(Accel, Speed_mean, Handling_mean, `M-turbo`,Traction)%>% skim_to_wide() %>% arrange(variable)

ggplot(summary, aes(W_category, median))+geom_point()+facet_grid(variable~., scales = "free")
```


```{r}
data_combo %>% filter(Speed_mean > 3 & Accel > 3 & Handling_mean > 3 & Speed_Land > 3 & W_category=="Medium") %>% top_n(10,total) %>% arrange(desc(total))
```

